---
title: "Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse')
library('caret')
library('gridExtra')
#install.packages('gridExtra')
raw_dat = ISLR::Auto %>%
  as_tibble() 

dat = raw_dat %>%
  select(acceleration, weight)

set.seed(123)
```

```{r}
raw_dat %>%
  ggplot(aes(x = weight, y = acceleration)) +
  geom_point()
```


Complete the following steps:  

  - Preprocess your data by one of the following (your choice): scale, center, standardize, normalize
  - Decide on an appropriate value of K ... via trial and error
  - Visualize your results in a plot
```{r}
summary(dat)
```

```{r}
dat %>%
  summarize(
    mean_weight = mean(weight),
    sd_weight = sd(weight),
    mean_acceleration = mean(acceleration),
    sd_acceleration = sd(acceleration))
```

```{r}
viz <- dat %>%
  ggplot(aes(x = weight, y = acceleration)) + 
  geom_point()
viz
```

```{r}
d1 <- dat %>%
  ggplot(aes(x = weight)) + 
  geom_density()

d2 <- dat %>%
  ggplot(aes(x = acceleration)) + 
  geom_density()

grid.arrange(d1, d2)
```

```{r}
result <- kmeans(dat, 5)
result
```

```{r}
result = kmeans(dat, 5)
viz2 <- dat %>%
  mutate(cluster = result$cluster) %>%
  ggplot(aes(x = weight, y = acceleration, col = as.factor(cluster))) + 
  geom_point()
gridExtra::grid.arrange(viz, viz2)
```

# ---- SCALING TRANSFORMATION---- ####  
## ---- Divides each value by the standard deviation of the column  

```{r}
scaled_func <- preProcess(dat, method = c("scale"))
scaled_dat <- predict(scaled_func, dat)
scaled_dat
```

```{r}
summary(scaled_dat)
```

```{r}
scaled_dat %>%
  summarize(
    mean_weight = mean(weight),
    sd_weight = sd(weight),
    mean_acceleration = mean(acceleration),
    sd_acceleration = sd(acceleration))
```

```{r}
sd_avg_weight <- mean(scaled_dat$weight)
sd_avg_accel <- mean(scaled_dat$acceleration)
sd_sd_weight <- sd(scaled_dat$weight)
sd_sd_accel <- sd(scaled_dat$acceleration)

sd1 <- scaled_dat %>%
  ggplot(aes(x = weight)) +
  geom_vline (xintercept = sd_avg_weight, color = 'red', linetype = 3, size = 1.2) +
  geom_vline (xintercept = sd_sd_weight, color = 'blue', linetype = 3, size = 1.2) +
  geom_density(fill="tan", alpha = .4)

sd2 <- scaled_dat %>%
  ggplot(aes(x = acceleration)) +
  geom_vline (xintercept = sd_avg_accel, color = 'red', linetype = 3, size = 1.2) +
  geom_vline (xintercept = sd_sd_accel, color = 'blue', linetype = 3, size = 1.2) +
  geom_density(fill="tan", alpha = .4)

gridExtra::grid.arrange(sd1, sd2)
```

```{r}
result = kmeans(scaled_dat, 5)
viz3 <- scaled_dat %>%
  mutate(cluster = result$cluster) %>%
  ggplot(aes(x = weight, y = acceleration, col = as.factor(cluster))) + 
  geom_point()
gridExtra::grid.arrange(viz, viz3)
```


# ---- CENTER TRANSFORMATION---- ####  
## ---- Subtracts the mean of each column from the values  

```{r}
centered_func = caret::preProcess(dat, method = c("center"))
centered_dat = predict(centered_func, dat)
centered_dat
```

```{r}
summary(centered_dat)
```

```{r}
centered_dat %>%
  summarize(
    mean_weight = mean(weight),
    sd_weight = sd(weight),
    mean_acceleration = mean(acceleration),
    sd_acceleration = sd(acceleration))
```

```{r}
cd_avg_weight <- mean(centered_dat$weight)
cd_avg_accel <- mean(centered_dat$acceleration)
cd_sd_weight <- sd(centered_dat$weight)
cd_sd_accel <- sd(centered_dat$acceleration)

cd1 <- centered_dat %>%
  ggplot(aes(x = weight)) +
  geom_vline (xintercept = cd_avg_weight, color = 'red', linetype = 3, size = 1.2) +
  geom_vline (xintercept = cd_sd_weight, color = 'blue', linetype = 3, size = 1.2) +
  geom_density(fill="tan", alpha = .4)

cd2 <- centered_dat %>%
  ggplot(aes(x = acceleration)) +
  geom_vline (xintercept = cd_avg_accel, color = 'red', linetype = 3, size = 1.2) +
  geom_vline (xintercept = cd_sd_accel, color = 'blue', linetype = 3, size = 1.2) +
  geom_density(fill="tan", alpha = .4)

gridExtra::grid.arrange(cd1, cd2)
```

```{r}
result = kmeans(centered_dat, 5)
viz4 <- centered_dat %>%
  mutate(cluster = result$cluster) %>%
  ggplot(aes(x = weight, y = acceleration, col = as.factor(cluster))) + 
  geom_point()
gridExtra::grid.arrange(viz, viz4)
```

# ---- STANDARDIZED TRANSFORMATION---- ####  
## ---- Devides each value by the column sd and subtrats the mean 

```{r}
standardized_func <- caret::preProcess(dat, method = c("center", "scale"))
standardized_dat <- predict(standardized_func, dat)
standardized_dat
```

```{r}
summary(standardized_dat)
```

```{r}
standardized_dat %>%
  summarize(
    mean_weight = mean(weight),
    sd_weight = sd(weight),
    mean_acceleration = mean(acceleration),
    sd_acceleration = sd(acceleration))
```

```{r}
std_avg_weight <- mean(standardized_dat$weight)
std_avg_accel <- mean(standardized_dat$acceleration)
std_sd_weight <- sd(standardized_dat$weight)
std_sd_accel <- sd(standardized_dat$acceleration)

std1 <- standardized_dat %>%
  ggplot(aes(x = weight)) +
  geom_vline (xintercept = std_avg_weight, color = 'red', linetype = 3, size = 1.2) +
  geom_vline (xintercept = std_sd_weight, color = 'blue', linetype = 3, size = 1.2) +
  geom_density(fill="tan", alpha = .4)

std2 <- standardized_dat %>%
  ggplot(aes(x = acceleration)) +
  geom_vline (xintercept = std_avg_accel, color = 'red', linetype = 3, size = 1.2) +
  geom_vline (xintercept = std_sd_accel, color = 'blue', linetype = 3, size = 1.2) +
  geom_density(fill="tan", alpha = .4)

gridExtra::grid.arrange(std1, std2)
```

```{r}
result = kmeans(standardized_dat, 5)
viz5 <- standardized_dat %>%
  mutate(cluster = result$cluster) %>%
  ggplot(aes(x = weight, y = acceleration, col = as.factor(cluster))) + 
  geom_point()
gridExtra::grid.arrange(viz, viz5)
```

# ---- NORMALIZED TRANSFORMATION ---- ####
## ---- Crunch everything between 0 and 1.  


```{r}
normalized_func <- caret::preProcess(dat, method = c("range"))
normalized_dat <- predict(normalized_func, dat)
normalized_dat
```

```{r}
summary(normalized_dat)
```

```{r}
normalized_dat %>%
  summarize(
    mean_weight = mean(weight),
    sd_weight = sd(weight),
    mean_acceleration = mean(acceleration),
    sd_acceleration = sd(acceleration))
```

```{r}
nd_avg_weight <- mean(normalized_dat$weight)
nd_avg_accel <- mean(normalized_dat$acceleration)
nd_sd_weight <- sd(normalized_dat$weight)
nd_sd_accel <- sd(normalized_dat$acceleration)

nd1 <- normalized_dat %>%
  ggplot(aes(x = weight)) +
  geom_vline (xintercept = nd_avg_weight, color = 'red', linetype = 3, size = 1.2) +
  geom_vline (xintercept = nd_sd_weight, color = 'blue', linetype = 3, size = 1.2) +
  geom_density(fill="tan", alpha = .4)

nd2 <- normalized_dat %>%
  ggplot(aes(x = acceleration)) +
  geom_vline (xintercept = nd_avg_accel, color = 'red', linetype = 3, size = 1.2) +
  geom_vline (xintercept = nd_sd_accel, color = 'blue', linetype = 3, size = 1.2) +
  geom_density(fill="tan", alpha = .4)

gridExtra::grid.arrange(nd1, nd2)
```

```{r}
result <- kmeans(normalized_dat, 5)
viz6 <- normalized_dat %>%
  mutate(cluster = result$cluster) %>%
  ggplot(aes(x = weight, y = acceleration, col = as.factor(cluster))) + 
  geom_point()
gridExtra::grid.arrange(viz,viz6)
```

## Running many times for each transformation.  

```{r}
k_val = 5

result_sd <- kmeans(scaled_dat, k_val)
p1 <- scaled_dat %>%
  mutate(cluster = result_sd$cluster) %>%
  ggplot(aes(x = weight, y = acceleration, col = as.factor(cluster))) + 
  geom_point() + 
  labs(title = 'scaled')

result_cd = kmeans(centered_dat, k_val)
p2 <- centered_dat %>%
  mutate(cluster = result_cd$cluster) %>%
  ggplot(aes(x = weight, y = acceleration, col = as.factor(cluster))) + 
  geom_point() + 
  labs(title = 'centered')

result_std = kmeans(standardized_dat, k_val)
p3 <- standardized_dat %>%
  mutate(cluster = result_std$cluster) %>%
  ggplot(aes(x = weight, y = acceleration, col = as.factor(cluster))) + 
  geom_point() + 
  labs(title = 'standardized')

result_nd = kmeans(normalized_dat, k_val)
p4 <- normalized_dat %>%
  mutate(cluster = result_nd$cluster) %>%
  ggplot(aes(x = weight, y = acceleration, col = as.factor(cluster))) + 
  geom_point() + 
  labs(title = 'normalized')

grid.arrange(p1, p2, p3, p4)
```


